---
title: "Statistical Inference Part 2"
author: "Nazmi Anik"
date: "Friday, October 23, 2015"
output: html_document
---

## Overview  

We are going to analyze the ToothGrowth data in the R datasets package. A basic summary of the data will be provided. We will use confidence intervals and/or hypothesis tests to compare tooth growth by supplement (vitamin c and ascorbic acid) and dose.  

## Load & Explore the Data  

We will load the data from the R package dataset.  

```{r load, echo=TRUE}
#load the data
data(ToothGrowth)
#do a quick summary to see what is inside
str(ToothGrowth)
```

Let's first load the libraries we will be using throughout our analysis.  

```{r library, message=FALSE}
#import ggplot andl dplyr libraries
library(ggplot2)
library(dplyr)
```

Now let's try to visualize the data with some boxplots. That will enable us to see the mean and variance in tooth growth with different dosages of each supplement.  

```{r boxplot1, echo=TRUE, fig.height=3.5, fig.width=5}
g <- ggplot(data = ToothGrowth, aes(x = factor(dose), y = len)) + geom_boxplot() + facet_grid(.~supp) + 
    xlab("Dose") + ylab("Tooth Length") + ggtitle("Tooth Growth with Different Supplements")
g
```

Here's another boxplot to see the mean and variance in tooth growth with different supplements for each dosage.
```{r boxplot2, echo=TRUE, fig.height=3.5, fig.width=5}
h <- ggplot(data = ToothGrowth, aes(x = factor(supp), y = len)) + geom_boxplot() + facet_grid(.~dose) + 
    xlab("Supplement") + ylab("Tooth Length") + ggtitle("Tooth Growth with Different Dosages")
h
```

It is clear from the plots that higher dosage of the supplements cause more tooth growth. The effects of using different supplements is less clear as their effects seems to be similar.  

## Basic Summary of Data  

Let's take a basic look at the data.  

```{r summary, echo=TRUE}
#group the data by dose and then by supplement
growth <- group_by(ToothGrowth, dose, supp)
#add mean and standard deviation to the summary table
growthSummary <- summarize(growth, mean = mean(len), standard.deviation = sd(len))
growthSummary
```

## Tooth Growth Comparison  

