---
title: "Statistical Inference Part 2"
author: "Nazmi Anik"
date: "Friday, October 23, 2015"
output: html_document
---

## Overview  

We are going to analyze the ToothGrowth data in the R datasets package. A basic summary of the data will be provided. We will use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.  

## Load the Data

We will load the data from the R package dataset.  

```{r load, echo=TRUE}
#load the data
data(ToothGrowth)
#do a quick summary to see what is inside
str(ToothGrowth)
```

Now let's try to visualize the data with some boxplots. That will enable us to see the mean and variance in tooth growth with different dosages of the different supplements.  

```{r, message=FALSE}
#import ggplot library
library(ggplot2)
```
```{r boxplot, echo=TRUE, fig.height=3.5, fig.width=5}
g <- ggplot(data = ToothGrowth, aes(x = factor(dose), y = len)) + geom_boxplot() + facet_grid(.~supp) + 
    xlab("Dose") + ylab("Tooth Length") + ggtitle("Tooth Growth with Different Supplements")
g
```

